
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 9: Visualization - Revealing a graph"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment9.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

**Notice:** *In this assignment, all the plot should have title, caption, and axis labels. *

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 

```{r}
library(gganimate)
library(gifski)
library(ggplot2)
library(tidyverse)
library(knitr)
library(lubridate)
df <- read_csv('https://covid19.who.int/WHO-COVID-19-global-data.csv') 
```


Hint: 

-  Use `filter` to select the most current date 
-  Pipe with `arrange(-Cumulative_deaths)` to rank the data by the number of cumulative deaths
```{r}
df %>%
  filter(Date_reported == '2021-12-31') %>%
  group_by(Country, Cumulative_deaths) %>%
  arrange(-Cumulative_deaths)

```



2. Use `transition_reveal` to make a plot revealing the number of deaths in the three countries with the most numbers of deaths
```{r}
df %>% 
  filter(Country_code %in% c('US','BR', 'IN')) %>% 
  ggplot(aes(y=Cumulative_deaths,
            x=Date_reported,
            color=Country_code))+ 
  geom_line()+
  geom_point(size=3)+
  geom_text(aes(label = Cumulative_deaths), 
            hjust = -.1, size=5) +
  transition_reveal(Date_reported)+
    labs(title = 'Number of Cumulative Deaths in Countries with Most Deaths', x='Date Reported', y='Cumulative Deaths', fill='Country', caption = 'Cumualative deaths US increase at a high rate than in Brazil or India from 2020 to 2023.')
```


3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 
```{r}
df$death_per_cases <- (df$Cumulative_deaths/df$Cumulative_cases)

df %>%
  filter(Date_reported == '2021-12-31') %>%
  group_by(Country) %>%
  arrange(-death_per_cases)
```


4. Use `transition_reveal` to make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 
```{r}
df %>% 
  filter(Country_code %in% c('US','IT', 'MX')) %>% 
  ggplot(aes(y=death_per_cases,
            x=Date_reported,
            color=Country_code))+ 
  geom_line()+
  geom_point(size=3)+
  geom_text(aes(label = death_per_cases), 
            hjust = -.1, size=5) +
  transition_reveal(Date_reported)+
    labs(title = 'Deaths Per Cases in US, IT, MX', x='Date Reported', y='Death Per Cases', fill='Country', caption = 'Death per cases is highest in Mexico compared to the USA and Italy from 2020 to 2023.')
```


5. Import the following data https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv
```{r}
df <- read_csv('https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv') 
```

                      
Use `transition_reveal` to plot the total vaccinations per hundred by level of income. Hint: the levels of income are identified in the `location` variable. Put a short comment on the caption of the plot. 

```{r}
df %>%
  filter(date >= '2023-05-01', date <= '2023-05-31', location %in% c('Low income','Lower middle income', 'Upper middle income', 'High income')) %>%
  ggplot(aes(y= total_vaccinations_per_hundred,
            x= date,
            color= location))+ 
  geom_line()+
  geom_point(size=3)+
  geom_text(aes(label = total_vaccinations_per_hundred), 
            hjust = -.1, size=5) +
  transition_reveal(date)+
    labs(title = 'Total Vaccinations per Hundred by Income Level', x='Date', y='Total Vaccinations per Hundred', fill='Location', caption = 'In May 2023, total vaccinations per hundred is greatest for high income individuals.' )
```

