
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 7 - Extra Credits: Practice dplyr and ggplot2"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment7_extra_credits.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas



```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```


-------

1. Import the data at https://bryantstats.github.io/math421/data/ds_salaries.csv

    The data description can be found at https://www.kaggle.com/datasets/saurabhshahane/data-science-jobs-salaries
    
```{r}
#Load packages.
library(tidyverse)
library(ggplot2)
library(ggthemes)

#read the data
df <- read_csv('ds_salaries.csv')
```


2. Is there any column that contains NAs?
```{r}
colSums(is.na(df))
#There are no columns that contain NAs.
```


3. What is the most frequent job title in the data?  

```{r}
df %>%
  group_by(job_title) %>%
  count() %>% 
  arrange(-n)

#The most frequent job title in the data is Data Scientist.
```


4. Find the average salary data of the three most frequent job titles in the data. 

```{r}
df %>%
  filter(job_title %in% c('Data Scientist', 'Data Engineer', 'Data Analyst')) %>%
  group_by(job_title) %>%
  summarise(averageSalaryUSD = mean(salary_in_usd))
```


5. What is the job title that have the highest average salary?
```{r}
df %>%
  group_by(job_title) %>%
  summarise(averageSalaryUSD = mean(salary_in_usd)) %>%
  arrange(-averageSalaryUSD)

#The job title that has the highest average salary in USD is Data Analytics Lead.
```


6. Which country gives the highest salary on average? 

```{r}
df %>%
  group_by(company_location) %>%
  summarise(averageSalaryUSD = mean(salary_in_usd)) %>%
  arrange(-averageSalaryUSD)

#The country that gives the highest salary on average is Russia (RU).
```


7. Plot five meaningful plots. All plots should have title, caption, and appropriate labels.
```{r}
df %>% 
  ggplot()+ 
  geom_bar(mapping=aes(x=company_size, fill=experience_level),
           position = 'dodge')+
labs(x='Company Size', 
       y = 'Number of Employees', 
       fill='Experience Level', 
       title = 'Distribution of Employee Experience Levels by Company Size ', 
       caption = 'Midsize have more senior level, large have more mid-level, small have about equal entry and mid level.')
```
```{r}
df %>%
  filter(job_title %in% c('Data Scientist', 'Data Engineer', 'Data Analyst')) %>%
  group_by(job_title) %>%
  summarise(averageSalaryUSD = mean(salary_in_usd)) %>%
  ggplot() + 
  geom_col(mapping = aes(x = job_title, y = averageSalaryUSD)) +
  labs(x = 'Job Title', 
       y = 'Average Salary in USD', 
       title = 'Average Salary by Top 3 Job Titles', 
       caption = 'The average salary in USD is highest for Data Engineers compared to Data Analysts or Data Scientists.')
```
```{r}
df %>%
  filter(company_location %in% c('US', 'CA', 'MX')) %>%
  group_by(company_location) %>%
  summarise(averageSalaryUSD = mean(salary_in_usd)) %>%
  ggplot() + 
  geom_col(mapping = aes(x = company_location, y = averageSalaryUSD)) +
  labs(x = 'Company Location', 
       y = 'Average Salary in USD', 
       title = 'Average Salary by North American Company Location', 
       caption = 'The average salary in USD for North American countries is highest in the US.')
```
```{r}
df %>%
  filter(company_location == 'MX')%>%
  group_by(work_year, company_location) %>%
  summarise(averageSalaryUSD = mean(salary_in_usd, na.rm = TRUE)) %>%
  ggplot() +
  geom_line(mapping = aes(x = work_year, 
                          y = averageSalaryUSD,
                          color = company_location)) +
  labs(x = 'Year', 
       y = 'Average Salary in USD', 
       color = 'Company Location', 
       title = 'Average Salary in Companies Located in Mexico from 2020-2022 ', 
       caption = 'Average salary in Mexican companies drops in 2021 followed by an increase from 2021 to 2022.')
```
```{r}
df %>%
  filter(company_location %in% c('GB', 'FR', 'GR'))%>%
  group_by(work_year, company_location) %>%
  summarise(averageSalaryUSD = mean(salary_in_usd, na.rm = TRUE)) %>%
  ggplot() +
  geom_line(mapping = aes(x = work_year, 
                          y = averageSalaryUSD,
                          color = company_location)) +
  labs(x = 'Year', 
       y = 'Average Salary in USD', 
       color = 'Company Location', 
       title = 'Average Salary in Companies Located in Europe from 2020-2022 ', 
       caption = 'When compared to GB and GR, FR has seen the most consistent increase in salary.')
```


