
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 3.  Working with Data Frame. Base R Style"
---

***Note*:** *This assignment practices working with Data Frame using Base R.*

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment3.Rmd)) in Rstudio. 

- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 3. 

-------

Problems
-------
 \

1. Create the following data frame

| Rank 	| Age 	| Name  	|
|:------	|:-----	|:-------	|
| 0    	| 28  	| Tom   	|
| 1    	| 34  	| Jack  	|
| 2    	| 29  	| Steve 	|
| 3    	| 42  	| Ricky 	|
```{r}
df = data.frame(Rank = c(0,1,2,3),
                Age = c(28, 34, 29, 42),
                Name = c("Tom", "Jack", "Steve","Ricky"))
df
```

2. Use `read.csv` to import the Covid19 Vaccination data from WHO: [link]((https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv)).
```{r}
df <- read.csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')
df
```

2. Show the names of the variables in the data
```{r}
names(df)
```

3. How many columns and rows the data have? 
```{r}
str(df)
#The data have 5 columns and 61,942 rows.
```

4. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?
```{r}
#There are no missing values.
sum(is.na(df))
```
```{r}
colSums(is.na(df))
```


```{r}
colMeans(is.na(df))
```

5. What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column. 
```{r}
class(df$date)
#The class of the data column is character.
```
```{r}
df$date = as.Date(df$date)
class(df$date)
```

6. Capitalize the names of all the variables
```{r}
names(df) <- c('Date', 'State', 'Flips', 'Cases', 'Deaths')
names(df)
```


7. Find the average number of cases per day. Find the maximum cases a day.
```{r}
mean(df$Cases, na.rm=TRUE)
#The average number of cases per day is 889830.1
```

```{r}
max(df$Cases, na.rm=TRUE)
#The maximum number of cases a day is 12169158
```
8. How many states are there in the data?  
```{r}
length(table(df$State))
#There are 56 states in the data.
```

9. Create a new variable `weekdays` to store the weekday for each rows.
```{r}
df$weekdays <- weekdays(df$Date)
df$weekdays
```

10. Create the categorical variable `death2` variable taking the values as follows

- `has_death` if there is a death that day
-  `no_death` if there is no death that day
```{r}
library(dplyr)
df$death2 <- case_when(df$Deaths==0 ~ 'no_death',
                       df$Deaths > 0 ~ 'has_death')
df$death2
```

Find the frequency and relative frequency of `no_death` and `has_death`.
```{r}
frequency <- table(df$death2)
relative_frequency <- prop.table(frequency)
#the frequency of has_death is 60790 and no_death is 1152
table(df$death2)
#the relative frequency of has_death is 0.898140196 and no_death is 0.01859804
prop.table(frequency)
```

11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint:  Use the `summary` function)
```{r}
summary(df$Deaths)
```

12. Create the categorical variable `death3` variable taking the values as follows

- `low_death` if the number of deaths smaller than the 25 percentile (Q1)

- `mid_death` if the number of deaths from Q1 to Q3

- `high_death` if the number of deaths greater than Q3
```{r}
#Obtain quartile data from prior summary function.
quartiles <- summary(df$Deaths)
Q1 <- quartiles["1st Qu."]
Q3 <- quartiles ["3rd Qu."]
#Create death3
df$death3 <- case_when(df$Deaths<Q1 ~ 'low_death',
                       df$Deaths<Q3 ~ 'mid_death',
                       TRUE~'high_death')
table(df$death3)
```

13. Find the average cases in Rhode Island in 2021
```{r}
df1 <- df[df$Date>='2021-01-01',]
by(df1$Cases, df1$State == 'Rhode Island', mean)
```


14. Find the median cases by weekdays in Rhode Island in 2021 
```{r}
df$weekdays <- weekdays(df$Date)
df1 = df[df$State=='Rhode Island',]
by(df1$Cases,df1$weekdays, median)
```

15. Compare the median cases in Rhode Island in June, July, August and September in 2021. 
```{r}
df$Date <- as.Date(df$Date)
df$YearMonth <- format(df$Date, "%Y-%m")
df1 <- df[df$State == 'Rhode Island' & df$YearMonth %in% c('2021-06', '2021-07', '2021-08', '2021-09'), ]
by(df1$Cases, df1$YearMonth, median)
```

16.  Find your own dataset, import it and implement the following functions on the data
```{r}
setwd("C:\\Users\\student\\OneDrive - Bryant University\\2023-2024\\Semester 1\\MATH421-A Statistical Analysis with R")
df <- read.csv('titanic.csv')
```

- head
```{r}
head(df)
```

- str
```{r}
str(df)
```

- names
```{r}
names(df)
```

- mean, min, max
```{r}
mean(df$Fare, na.rm = TRUE)
min(df$Fare, na.rm = TRUE)
max(df$Fare, na.rm = TRUE)
```

- table
```{r}
table(df$Pclass)
```

- is.na
```{r}
sum(is.na(df))
```

- colSums
```{r}
colSums(is.na(df))
```

- class
```{r}
class(df$Survived)
```

- cor
```{r}
cor(df$Survived, df$Pclass)
```

- by
```{r}
by(df$Fare, df$Ticket, FUN = mean)
```

- ifelse
```{r}
df$Survived2 <- ifelse(df$Survived==0, 
                        'no_survivors','has_survivors')
```

- case_when
```{r}
library(dplyr)
df$Survived2 <- case_when(df$Survived==0 ~ 'no_survivors',
                       df$Survived > 0 ~ 'has_survivors')
df$Survived2
```



17.  In the dataset in #16, practice the follows.  You can reuse the code of 16. 

- Create a categorical variable from a continuous variable
```{r}
library(dplyr)
df$Age2 <- case_when(df$Age < 18 ~ 'child',
                       df$Age > 18 ~ 'adult')
df$Age2
```

- From one categorical variable, create a new categorical variable with fewer categories
```{r}
library(dplyr)
df$Embarked2 <- case_when(df$Embarked %in% c('Q', 'C') ~ 'Europe',
                          df$Embarked %in% c('S') ~ 'England')
df$Embarked2
```


If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](../data/titanic.csv)
