
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 7: Visualization - Some aspects of ggplot2"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment7.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

-------

Use the data of your own.  Produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. All plots should have title, caption, appropriate labels on x and y-axis. 

-------
```{r}
#Load packages.
library(tidyverse)
library(ggplot2)
library(ggthemes)

#read the data
df <- read_csv('adult_census.csv')
```


1. Make a plot using `facet_wrap` by one categorical variable (`facet_wrap(~variable1)`)
```{r}
df %>% 
  filter(sex %in% c("Male", "Female")) %>%
  ggplot()+ 
  geom_point(mapping=aes(x=age, y=fnlwgt))+
  facet_wrap(~sex)+
labs(x='Age', 
       y = 'Final Weight', 
       title = 'Final Weight of Census Genders by Age', 
       caption = 'As age increases for both reported sexes in the census, the lower the final weight. This suggests the number of people the census believes the entry represents decreases.')
```
2. Make a plot using `facet_wrap` by two categorical variables (`facet_wrap(~variable1+variable2)`)
```{r}
df %>% 
  filter(sex %in% c("Male", "Female")) %>%
  ggplot()+ 
  geom_point(mapping=aes(x=age, y=fnlwgt))+
  facet_wrap(~sex+income)+
labs(x='Age', 
       y = 'Final Weight', 
       title = 'Final Weight of Census Genders and Incomes by Age ', 
       caption = 'As age increases for both reported sexes and incomes in the census, the lower the final weight suggesting the number of people the census believes the entry represents decreases.')
```


3. Make a plot using `position ='dodge'`.
```{r}
df %>% 
  filter(sex %in% c("Male", "Female")) %>%
  ggplot()+ 
  geom_bar(mapping=aes(x=sex, fill=race),
           position = 'dodge')+
labs(x='Sex', 
       y = 'Number of Census Respondents', 
       fill='Race', 
       title = 'Number of Races Represented in Census by Gender', 
       caption = 'White males have the highest frequency of all census respondents followed by white females.')
```


4. Make a plot using `position ='fill'`. 
```{r}
df %>% 
  filter(sex %in% c("Male", "Female")) %>%
  ggplot()+ 
  geom_bar(mapping=aes(x=sex, fill=income), 
           position = 'fill')+
  labs(x='Sex', 
       y = 'Percentage of Census Respondents', 
       fill='Income', 
       title = 'Income of Census Respondents by Gender', 
       caption = 'Both male and female respondents have a greater percentage of income <=50K than >50K in this census.')
  

```


5. Make a plot using `geom_col`.
```{r}
df %>%
  filter(sex %in% c("Male", "Female")) %>%
  group_by(sex) %>% 
  summarise(average_age=mean(age, na.rm=TRUE)) %>% 
  ggplot()+ 
  geom_col(mapping=aes(x=sex, y=average_age))+
  labs(x='Sex', 
       y = 'Average Age (Years)', 
       title = 'Average Ages of Census Respondents by Gender', 
       caption = 'The average age of females in this census is approximately 36-37 while the average age of males is approximately 39 years .')
```


6. Change the theme of one of your plots.
```{r}
df %>% 
  filter(sex %in% c("Male", "Female")) %>%
  ggplot()+ 
  geom_bar(mapping=aes(x=sex, fill=race),
           position = 'dodge')+
  theme_dark()+
labs(x='Sex', 
       y = 'Number of Census Respondents', 
       fill='Race', 
       title = 'Number of Races Represented in Census by Gender', 
       caption = 'White males have the highest frequency of all census respondents followed by white females.')
```


7. Make a plot using combo filter + ggplot
```{r}
df %>% 
    filter(age>=25, age<=50) %>% 
    ggplot()+
    geom_bar(mapping=aes(x=race, fill=income))+
labs(x='Race', 
       y = 'Number of Census Respondents', 
       fill='Race', 
       title = 'Income by Race in Census Data', 
       caption = 'People who identify as white make up the greatest frequency in this data. All races in this dataset primarily make <=50K income rather than >50K.')
```


8. Make a plot using combo group_by + summarise + geom_col
```{r}
df %>% 
  filter(sex %in% c("Male", "Female")) %>%
  group_by(income, sex) %>% 
  summarise(mean_educationNum = education.num) %>% 
  ggplot()+
  geom_col(aes(x=sex, y=mean_educationNum, fill=income), position = 'dodge')+
labs(x='Race', 
       y = 'Average Education Years by Gender and Income', 
       fill='Income', 
       title = 'Income by Gender', 
       caption = 'Interestingly, when grouping by both gender and income, the average years of education for all groups is approximately equal and about 17 years.')
```


9. Save a plot and put a link of the plot to your Github Webpage.  To link a plot `abc.png` you just need to insert the following on the README.md

```{r, eval=FALSE}
#[Link to the plot](abc.png)

gg <- df %>% 
  filter(sex %in% c("Male", "Female")) %>%
  ggplot()+ 
  geom_bar(mapping=aes(x=sex, fill=race),
           position = 'dodge')+
labs(x='Sex', 
       y = 'Number of Census Respondents', 
       fill='Race', 
       title = 'Number of Races Represented in Census by Gender', 
       caption = 'White males have the highest frequency of all census respondents followed by white females.')

ggsave(filename = 'Race_by_Gender.png', plot = gg)
```

10. In this question, we will use  the Covid19 vaccination data for the US. The data and data dictionary can be found at the below link:

- [Data](https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations-by-manufacturer.csv)

```{r}
library(tidyverse)
df <- read_csv('https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations-by-manufacturer.csv')
```

`Geom_point` makes scatter plots, which plot two numeric (or Date) variables. A third categorical variable is often brought in for color. For example, we can compare different vaccine types used in the US by

```{r}
df %>% 
  filter(location=='United States') %>% 
  ggplot()+
  geom_point(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine))+
labs(x='Date', 
       y = 'Total Vaccinations', 
       fill='Vaccine', 
       title = 'Vaccines Used in United States by Date', 
       caption = 'Throughout the specified time period, the Pfizer/BioNTech vaccine is most frequently used in the USA.')
```

Do the follows. 

  - Use `geom_point` to compare the vaccines used by date in Argentina in 2022. 
```{r}
dateFormatted <- as.Date(df$date)

df %>% 
  filter(location=='Argentina', date>="2022-01-01",date<"2023-01-01") %>%
  ggplot()+
  geom_point(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine))+
labs(x='Date', 
       y = 'Total Vaccines', 
       fill='Vaccines', 
       title = 'Vaccines Used in Argentina in 2022', 
       caption = 'The Sinopharm/Beijing vaccine was the most commonly used vaccine in Argentina in 2022.')

```

  
  - Change `geom_point` to `geom_line` in the above codes to have the corresponding line plots. 
```{r}
#USA 
df %>% 
  filter(location=='United States') %>% 
  ggplot()+
  geom_line(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine))+
labs(x='Date', 
       y = 'Total Vaccinations', 
       fill='Vaccine', 
       title = 'Vaccines Used in United States by Date', 
       caption = 'Throughout the specified time period, the Pfizer/BioNTech vaccine is most frequently used in the USA.')

#Argentina 
df %>% 
  filter(location=='Argentina', date>="2022-01-01",date<"2023-01-01") %>%
  ggplot()+
  geom_line(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine))+
labs(x='Date', 
       y = 'Total Vaccines', 
       fill='Vaccines', 
       title = 'Vaccines Used in Argentina in 2022', 
       caption = 'The Sinopharm/Beijing vaccine was the most commonly used vaccine in Argentina in 2022.')

```


11. (Using the same dataset as in 10.) Sometime we need to do some calculations on the data before plotting. For example, the below codes is to compare the total vaccines (including all types of vaccines) by date between the US and the EU

```{r}
# the total vaccines (including all types of vaccines) by date
d1 <- df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE))

# Plot
d1 %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))+
labs(x='Date', 
       y = 'Total Vaccinations', 
       fill='Vaccine', 
       title = 'Vaccines Used in European Union and United States by Date', 
       caption = 'Throughout the specified time period, the European Union used a greater total number of vaccinations compared to the USA.')
```

Notice that the above codes could be piped in one 

```{r, eval=FALSE}
df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac=2sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))+
labs(x='Date', 
       y = 'Total Vaccinations', 
       fill='Vaccine', 
       title = 'Vaccines Used in United States and  European Union by Date', 
       caption = 'Throughout the specified time period, the European Union used a greater total number of vaccinations compared to the USA.')
```


Do the follows. 

- Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in Asia
```{r}

df %>%
  filter(location %in% c('South Korea','Hong Kong', 'Japan') ) %>%
  group_by(date, location) %>%
  summarise(total_vac2 = sum(total_vaccinations, na.rm = TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))+
labs(x='Date', 
       y = 'Total Vaccinations', 
       fill='Vaccine', 
       title = 'Vaccines Used in Asian Countries and Regions by Date', 
       caption = 'Of the countries and regions from Asia included in this dataset, Japan uses the greatest total number of vaccines compared to Hong Kong and South Korea.')
```


- Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in South America in 2021
```{r}


df %>% 
  filter(location %in% c('Argentina', 'Chile', 'Ecuador', 'Peru', 'Portugal', 'Spain', 'Uruguay'), date>="2021-01-01",date<"2022-01-01") %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))+
labs(x='Date', 
       y = 'Total Vaccinations', 
       fill='Vaccine', 
       title = 'Vaccines Used in South American Countries and Regions in 2021', 
       caption = 'Of the South American countries and regions in this dataset, Spain uses the most total vaccinations in 2021.')
```


