
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 15: Writing a function - Conditional Statement"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment15.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

#### NOTICE: Question 7, 8 and 9 are optional. 

-------

1. Write a function to calculate the age of a person.  Give an example to test your function. 

- Input: a number, x, (year born)

- Output: Print out "You are r age". Where r is the age of the person, i.e. 2020 - x. 

Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#42)

```{r}
age <- function(x)
{
  r = 2023 - x 
  print(paste0('You are ', r, ' years old'))
}
```

```{r}
age(2002)
```

-------

2. Write the following function. Give an example to test your function. 

- Input: a number

- Output: print out: "You input an even number!" if the number is event, or "You input an odd number!" otherwise. 

Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#46)

```{r}
numberType <- function(x)
{
  if(x %% 2 == 0)
  {
    print('You input an even number!') 
  }
  else
  {
    print('You input an odd number!')
  }
}
```

```{r}
#Input an even number
numberType(14)
```
```{r}
#Input an odd number
numberType(51)
```

-------

3. Write the following function.  Give an example to test your function. 

- Input: a numeric vector 

- Output:  

  - if the input vector has missing values: return the input vector with missing values replaced by mean
  
  - if the input vector has no missing value: return the same input vector
  
Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#50)

```{r}
mean_input <- function(x) {
  if (any(is.na(x))) {
    if (is.numeric(x)) {
      x <- ifelse(is.na(x), mean(x, na.rm = TRUE), x)
    }
  }
  return(x)
}
  
```

```{r}
#Numeric vector with missing data
x1 <- c(1, 2, 3, NA, 5)
mean_input(x1)
```
```{r}
#Complete numeric vector
x2 <- c(1, 2, 3, 4, 5)
mean_input(x2)
```

-------

4. Write the following function.  Give an example to test your function. 

- Input: a vector x 

- Output:  The vector x where the missing values replaced by the mean (if x is numeric) or the mode (if x is non-numeric).  If x does not have missing value, return the same vector x.  

Hint: Use If-statement to combine the function in Question 3 and [this function](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#49)

```{r}
vector_input <- function(x) {
  if (any(is.na(x))) {
    if (is.numeric(x)) {
      x <- ifelse(is.na(x), mean(x, na.rm = TRUE), x)
    } else {
      mode_val <- names(sort(table(x), decreasing = TRUE)[1])
      x <- ifelse(is.na(x), mode_val, x)
    }
  }
  return(x)
}
```

```{r}
#Numeric vector with missing data
x <- c(1, 2, 3, NA, 5)
vector_input(x)
```
```{r}
#Non-numeric data with missing data
n <- c("a", "a", "b", "c", "d", NA)
vector_input(n)
```
```{r}
#Complete numeric vector
m <- c(10, 20, 30, 40, 50)
vector_input(m)
```

-------

5. Write the following function. Give examples to test your function. 

- Input: A data frame of two variables x and y

- Output: 

  - A boxplot of x by y if x is numeric and y is non-numeric
 
  - A boxplot of y by x if y is numeric and x is non-numeric
 
  - print out 'This function cannot visualize your data' otherwise

Hint: 

- [Similar Function](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#53)
 
- You can refer to this slide to plot a boxplot: https://bryantstats.github.io/math421/slides/6_viz.html#36

```{r}
viz_boxplot <- function(data) {
  library(ggplot2)
  if (is.numeric(data[[1]]) && !is.numeric(data[[2]])) {
    ggplot(data, aes(x = data[[2]], y = data[[1]])) +
      geom_boxplot() +
      labs(x = names(data)[2], y = names(data)[1]) +
      ggtitle("Boxplot of Numeric Variable by Non-Numeric Variable")
  } else if (is.numeric(data[[2]]) && !is.numeric(data[[1]])) {
    ggplot(data, aes(x = data[[1]], y = data[[2]])) +
      geom_boxplot() +
      labs(x = names(data)[1], y = names(data)[2]) +
      ggtitle("Boxplot of Numeric Variable by Non-Numeric Variable")
  } else {
    print("This function cannot visualize your data.")
  }
}
```

```{r}
#X is numeric, Y is not numeric
a <- data.frame(x = rnorm(100), y = rep(c("A", "B"), each = 50))
viz_boxplot(a)
```
 
```{r}
#X is not numeric, Y is numeric
b <- data.frame(x = rep(c("A", "B"), each = 50), y = rnorm(100))
viz_boxplot(b)
```

```{r}
#Data that cannot be visualized by function
c <- data.frame(x = rnorm(100), y = rnorm(100))
viz_boxplot(c)
```

-------
 
6. Combine the function in Question 5 and the function in this example ([Link](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#53)) to have a function that can plot all possible cases (2 categorical, 2 numeric, 1 numeric and 1 categorical). 

```{r}
new_combined_viz <- function(data) {
  if (is.numeric(data[[1]]) && is.numeric(data[[2]])) {
    ggplot(data, aes_string(x = names(data)[1], y = names(data)[2])) +
      geom_point() +
      labs(x = names(data)[1], y = names(data)[2]) +
      ggtitle("Scatter Plot of Numeric Variables")
  } else if (!(is.numeric(data[[1]]) | is.numeric(data[[2]]))) {
    ggplot(data, aes_string(x = names(data)[1], fill = names(data)[2])) +
      geom_bar(position = 'dodge') +
      labs(x = names(data)[1], fill = names(data)[2]) +
      ggtitle("Grouped Bar Plot of Categorical Variables")
  } else {
    ggplot(data, aes_string(x = names(data)[1], y = names(data)[2])) +
      geom_boxplot() +
      labs(x = names(data)[1], y = names(data)[2]) +
      ggtitle("Boxplot of Numeric Variable by Non-Numeric Variable")
  }
}
```

```{r}
#X is numeric, Y is not numeric
x <- data.frame(x = rnorm(100), y = rep(c("A", "B"), each = 50))
new_combined_viz(x)
```
```{r}
#X is not numeric, Y is numeric
x <- data.frame(x = rep(c("A", "B"), each = 50), y = rnorm(100))
new_combined_viz(x)
```

```{r}
#X and Y are both numeric

n <- data.frame(x = rnorm(100), y = rnorm(100))
new_combined_viz(n)

```
```{r}
#Both x and y are non-numeric
b <- data.frame(x = rep(c("A", "B"), each = 50), y = rep(c("C", "D"), each = 50))
new_combined_viz(b)

```

-------

#### NOTICE: Question 7, 8 and 9 are optional. 

-------

7. Write the following function. Give examples to test your function. 

- Input: 

  - input_data: a clean data frame with a variable name `target`. The `target` variable is also binary. 
  
  - train_percent: a number presenting a proportion of training data. The default train_percent is .8

- Output: the accuracy of the decision model `rpart` where the training data is train_percent. 

```{r}
modl <- function(input_data, train_percent= .8)
{
library(caret)
library(tidyverse)
library(tidyr)
library(dplyr)
set.seed(00000)
splitIndex <- createDataPartition(input_data$target, p = train_percent, 
                                  list = FALSE)
df_train <- input_data[ splitIndex,]
df_test <- input_data[-splitIndex,]
library(rpart)
tree1<-rpart(target ~ ., 
             data = df_train)
# Plot the tree
library(rattle)
fancyRpartPlot(tree1)
}
```

```{r}
# Read in the data
library(tidyverse)
df = read_csv("titanic.csv")
# Set the target variable
names(df)[8] <- 'target'
# Remove some columns
df$PassengerId =  NULL
df$Ticket =  NULL
df$Name = NULL
df$Cabin = NULL
# Correct variables' types
df$target <- factor(df$target)
df$Pclass = factor(df$Pclass)
df$Sex <- factor(df$Sex)
df$Embarked <- factor(df$Embarked)
# Handle missing values
df$Age[is.na(df$Age)] = mean(df$Age, na.rm = TRUE)
df = drop_na(df)
```

```{r}
#Test decision tree function
modl(df)
```

-------

8. Write the following function. Give examples to test your function. 

- Input: 

  - input_data: a clean data frame with a variable name `target`. The `target` variable is also binary. 
  
  - train_percent: a number presenting a proportion of training data. The default train_percent is .8

- Output: the plot of variable important by random forest trained by caret. 

```{r}
modl2 <- function(input_data, train_percent=.8)
{
  library(caret)
  library(dplyr)
  library(tidyverse)
  library(readr)
  library(tidyr)
set.seed(00000)
splitIndex <- createDataPartition(input_data$target, p = train_percent, 
                                  list = FALSE)
df_train <- input_data[ splitIndex,]
df_test <- input_data[-splitIndex,]
model <- train(target~., data=df_train, 
                        method = "ranger")
pred <- predict(model, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
}
```

```{r}
# Read in the data
df = read_csv("titanic.csv")
# Set the target variable
names(df)[8] <- 'target'
# Remove some columns
df$PassengerId =  NULL
df$Ticket =  NULL
df$Name = NULL
df$Cabin = NULL
# Correct variables' types
df$target <- factor(df$target)
df$Pclass = factor(df$Pclass)
df$Sex <- factor(df$Sex)
df$Embarked <- factor(df$Embarked)
# Handle missing values
df$Age[is.na(df$Age)] = mean(df$Age, na.rm = TRUE)
df = drop_na(df)
```

```{r}
#Test function 
suppressWarnings(modl2(df))
```

---
9. Write the following function. Give examples to test your function. 

- Input: 

  - a data frame that has a text column 
  
  - the name of the text column in the data

- Output: the word cloud plot of the text column

Sample codes

- [Word Cloud codes](https://bryantstats.github.io/math421/assignments/text_mining_sample_codes2.html#Total_Word_Cloud)

- [A similar function](https://bryantstats.github.io/math421/gganimate/17_writing_functions.html#10)

```{r}
word_cloud <- function(text_df, text_col) {
  library(tidyverse)
  library(tidytext)
  library(wordcloud)
  pal <- brewer.pal(8,"Dark2")
  df %>%
  unnest_tokens(input = description, output = word) %>% 
  anti_join(get_stopwords()) %>% 
  count(word, sort = TRUE) %>%
  with(wordcloud(word, n, random.order = FALSE, max.words = 50, colors=pal))
}
```

```{r}
df <- read_csv('netflix_titles.csv')
word_cloud(df)
```

