
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 10: Predictive Modeling - Part 1"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment10.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Use the `Adult Census Income` dataset.  We will predict the income (whether or not it is more than 50k or not) of an adult. Import the dataset.  Partition the data into 80% training and 20% testing.  

```{r}
library(tidyverse)
df <- read_csv('adult_census.csv')
```

```{r}
#take out some columns 
df <- df %>% select(-fnlwgt, -marital.status, -occupation, -relationship, -capital.gain, -capital.loss)

#set the target variable 
df <- df %>% rename(target=income)
```

```{r}
#changing character variables to factors 
df <- df %>% 
  mutate(target = as.factor(target),
         workclass = as.factor(workclass),
         race = as.factor(race),
         sex = as.factor(sex),
         education = as.factor(education),
         native.country = as.factor(native.country)
         )

#drop NA values 
df = drop_na(df)
```

```{r}
#install.packages("caret")
library(caret)
set.seed(2023)
splitIndex <- createDataPartition(df$target, p = .80, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```

2. Practice Decision Tree.  Do the follows:

  - Use `rpart` package, create a decision tree with maximum depth of 3. 
```{r}
#install.packages("rpart")
library(rpart)

decision_tree <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 3))
```
  
  
  - Calculate the accuracy of the model on the testing data. Notice that the positive outcome here is not `1` but `>50K` or `<50K`. 
```{r}
pred <- predict(decision_tree, df_test, type = "class")
#Evaluate the predictions
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = ">50K")
cm$overall[1]
```

  
  - Plot the tree
```{r}
#install.packages("rattle")
library(rattle)
fancyRpartPlot(decision_tree)
```

  
  - Plot the variable importance by the tree
```{r}
decision_tree$variable.importance
```

```{r}
barplot(decision_tree$variable.importance)
```
  
3. Create 3 more trees and compare the testing accuracy of these trees, which tree give the highest testing accuracy.
```{r}
set.seed(2010)
splitIndex <- createDataPartition(df$target, p = .80, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]

tree1 <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 11))

fancyRpartPlot(tree1)
```

```{r}
#accuracy of tree1 
pred <- predict(tree1, df_test, type = "class")
#Evaluate the predictions
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = ">50K")
cm$overall[1]
```


```{r}
set.seed(1000)
splitIndex <- createDataPartition(df$target, p = .80, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]

tree2 <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 30))

fancyRpartPlot(tree2)
```

```{r}
#accuracy of tree2
pred <- predict(tree2, df_test, type = "class")
#Evaluate the predictions
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = ">50K")
cm$overall[1]

#Of the trees trained, tree 2 yields the highest accuracy.
```

```{r}
set.seed(1780)
splitIndex <- createDataPartition(df$target, p = .80, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]

tree3 <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 6))

fancyRpartPlot(tree3)
```

```{r}
#accuracy of tree3
pred <- predict(tree3, df_test, type = "class")
#Evaluate the predictions
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = ">50K")
cm$overall[1]
```

4. Practice Random Forest.  Do the follows: 

  - Use `randomForest` package, create a random forest of 1000 trees. 
  
```{r}
#install.packages("randomForest")
library(randomForest)
forest_model = randomForest(target ~ ., data=df_train, ntree = 1000)
pred <- predict(forest_model, df_test, type = "class")
```
  
  
  - Calculate the accuracy of the model on the testing data. 
```{r}
cm <- confusionMatrix(data = pred, reference = df_test$target, positive = ">50K")
cm$overall[1]
```
  
  
  - Plot the variable importance by the forest
```{r}
importance(forest_model)
```
  

5. Create 3 more forests and compare the testing accuracy of these forests, which forest give the highest testing accuracy.

```{r}
forest_model1 = randomForest(target ~ ., data=df_train, ntree = 2023)
pred <- predict(forest_model1, df_test, type = "class")

cm <- confusionMatrix(data = pred, reference = df_test$target, positive = ">50K")
cm$overall[1]

#of the forest trained, forest_model1 yields the highest testing accuracy 
```

```{r}
forest_model2 = randomForest(target ~ ., data=df_train, ntree = 50)
pred <- predict(forest_model2, df_test, type = "class")

cm <- confusionMatrix(data = pred, reference = df_test$target, positive = ">50K", "<50K")
cm$overall[1]
```


```{r}
forest_model3 = randomForest(target ~ ., data=df_train, ntree = 4000)
pred <- predict(forest_model3, df_test, type = "class")

cm <- confusionMatrix(data = pred, reference = df_test$target, positive = ">50K")
cm$overall[1]
```

6. What is the best model (in term of testing accuracy) among all models (including trees and forests) you have trained?

```{r}
library(randomForest)
forest_model = randomForest(target ~ ., data=df_train, ntree = 1000)
pred <- predict(forest_model, df_test, type = "class")


#Of all models, I've trained, this model yields the highest testing accuracy. Accuracy :0.8126059 
```

