
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 16: Writing a function - Loop"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment16.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

1. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: a data frame with all the missing of numeric variables replaced by the associated means.

Hint: [Similar function](https://bryantstats.github.io/math421/gganimate/18_writing_functions.html#28)

```{r}
numeric_impute <- function(d) {
  for (i in 1:ncol(d)) {
    if (is.numeric(d[[i]])) {
      d[[i]][is.na(d[[i]])] <- mean(d[[i]], na.rm = TRUE)
    }
  }
  return(d)
}
```

```{r}
#Test numeric_impute function.
library(tidyverse)
df  <- read_csv('adult_census.csv')

colSums(is.na(df))
```
```{r}
df1 <- numeric_impute(df)
colSums(is.na(df1))
```


-------

2. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: a data frame with all the missing of variables replaced by the associated means (for numeric variables) or modes (for non-numeric variables).

Hint: Use If-statement to combine the function in Problem 1 and the function in [this example](https://bryantstats.github.io/math421/gganimate/18_writing_functions.html#28)

```{r}
impute_na <- function(d) {
  for (i in 1:ncol(d)) {
    if (is.numeric(d[[i]])) {
      d[[i]][is.na(d[[i]])] <- mean(d[[i]], na.rm = TRUE)
    } else if (any(!is.na(d[[i]]))) {
      levels <- unique(d[[i]])
      mode_val <- levels[which.max(tabulate(match(d[[i]], levels)))]
      d[[i]][is.na(d[[i]])] <- mode_val
    }
  }
  return(d)
}
```

```{r}
#Test impute_na function.
library(tidyverse)
df  <- read_csv('adult_census.csv')

colSums(is.na(df))
```
```{r}
dfNew <- impute_na(df)
colSums(is.na(dfNew))
```

-------

3. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: Bar plots of all non-numeric variables

Hint: [Similar function](https://bryantstats.github.io/math421/gganimate/18_writing_functions.html#33)

```{r}
bar_plot <- function(d) {
  library(ggplot2)
  for (i in 1:ncol(d)) {
    if (!is.numeric(d[[i]])) {
      print(ggplot(d, aes_string(x = names(d)[i])) +
            geom_bar() +
            labs(x = names(d)[i]))
    }
  }
}
```

```{r}
#Test bar_plot function.
library(tidyverse)
df <- read_csv('adult_census.csv')

bar_plot(df)
```
```{r}
#Test bar_plot function.
library(tidyverse)
df <- read_csv('adult_census.csv')

bar_plot(df)
```

-------

4. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: all possible the bar plots of a non-numeric variable filled by a non-numeric variable.

Hint: [Similar function](https://bryantstats.github.io/math421/gganimate/18_writing_functions.html#40)

```{r}
bar_plots_filled <- function(d) {
  library(ggplot2)
  non_numeric_vars <- sapply(d, function(x) !is.numeric(x))
  non_numeric_cols <- names(d)[non_numeric_vars]
  
  for (col1 in non_numeric_cols) {
    for (col2 in non_numeric_cols) {
      if (col1 != col2) {
        plot_data <- d %>%
          group_by(!!sym(col1), !!sym(col2)) %>%
          summarise(count = n())
        
        print(ggplot(plot_data, aes(x = !!sym(col1), y = count, fill = !!sym(col2))) +
                geom_bar(position = "dodge", stat = "identity") +
                labs(x = col1, y = "Count", fill = col2) +
                theme_minimal())
      }
    }
  }
}

```

```{r}
#Test the bar_plots_filled function

df <- read_csv('adult_census.csv')

bar_plots_filled(df)
```

-------

5. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: 

  - all possible the bar plots of a non-numeric variable filled by a non-numeric variable.
  
  - all possible the density plots of a numeric variable colored by a non-numeric variable
  
  - all possible the scatter plots.

Hint: Combine [this function](https://bryantstats.github.io/math421/gganimate/18_writing_functions.html#37), [this function](https://bryantstats.github.io/math421/gganimate/18_writing_functions.html#40), and the function in Question 4. One way to combine is creating a new function, `quick_plot`, and call these three functions within `quick_plot`. 

```{r}
library(ggplot2)
library(dplyr)

bar_plots_filled <- function(d) {
  library(ggplot2)
  non_numeric_vars <- sapply(d, function(x) !is.numeric(x))
  non_numeric_cols <- names(d)[non_numeric_vars]
  
  for (col1 in non_numeric_cols) {
    for (col2 in non_numeric_cols) {
      if (col1 != col2) {
        plot_data <- d %>%
          group_by(!!sym(col1), !!sym(col2)) %>%
          summarise(count = n())
        
        print(ggplot(plot_data, aes(x = !!sym(col1), y = count, fill = !!sym(col2))) +
                geom_bar(position = "dodge", stat = "identity") +
                labs(x = col1, y = "Count", fill = col2) +
                theme_minimal())
      }
    }
  }
}


# Example usage:
# bar_plots_filled(your_data_frame)


# Example usage:
# bar_plots_filled(your_data_frame)


#Density Plots
density_plots <- function(d) {
  numeric_vars <- sapply(d, is.numeric)
  numeric_cols <- names(d)[numeric_vars]
  non_numeric_vars <- sapply(d, function(x) !is.numeric(x))
  non_numeric_cols <- names(d)[non_numeric_vars]
  
  for (num_col in numeric_cols) {
    for (non_num_col in non_numeric_cols) {
      print(ggplot(d, aes(x = !!sym(num_col), fill = !!sym(non_num_col))) +
              geom_density(alpha = 0.5) +
              labs(x = num_col, fill = non_num_col) +
              theme_minimal())
    }
  }
}

#Scatterplots
scatter_plots <- function(d) {
  numeric_vars <- sapply(d, is.numeric)
  numeric_cols <- names(d)[numeric_vars]
  non_numeric_vars <- sapply(d, function(x) !is.numeric(x))
  non_numeric_cols <- names(d)[non_numeric_vars]
  
  for (num_col1 in numeric_cols) {
    for (num_col2 in numeric_cols) {
      if (num_col1 != num_col2) {
        print(ggplot(d, aes(x = !!sym(num_col1), y = !!sym(num_col2))) +
                geom_point() +
                labs(x = num_col1, y = num_col2) +
                theme_minimal())
      }
    }
  }
}

#Function quick_plot
quick_plot <- function(d) {
  bar_plots_filled(d)
  density_plots(d)
  scatter_plots(d)
}
```

```{r}
#Test quick_plot function

df <- read_csv('adult_census.csv')

suppressWarnings(quick_plot(df))
```




