---
title: "MATH 421 Midterm Project"
author: "Katelyn Caldarone"
date: "2023-10-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Use read_sas in library haven to read the hospital discharge data. 
```{r}
library(haven)
data <- read_sas('hdd0318cy.sas7bdat')
```

2. Filter the data to have only patients of the year 2018(yod = 2018)
```{r}
library(tidyverse)

data %>%
  filter(yod == 18)
```

3. Select to work only with the following variables:
```{r}
workingVariables <- c("yod", "payfix","pay_ub92","age",  
                      "sex","raceethn","provider","moa", 
                      "yoa","mod","admtype", "asource" , 
                      "preopday" ,"los", "service" , "icu","ccu",    
                      "dispub92", "payer"  ,"drg","trandb", 
                      "randbg","randbs","orr", "anes","seq",   
                      "lab","dtest", "ther","blood","phar", 
                      "other","patcon","bwght","total","tot" ,  
                      "ecodub92","b_wt","pt_state","diag_adm","ancilar" ,
                      "campus","er_fee","er_chrg","er_mode","obs_chrg",
                      "obs_hour","psycchrg","nicu_day")

dataReduced <- data[workingVariables]
```

4.What are the variables with missing data?
```{r}

df <- read_csv('midterm.csv', show_col_types = FALSE)
```
```{r}
df <- replace(df, df == ' ', NA)

colSums(is.na(df))
```


5. Remove all variables with missing variables.
```{r}
library(dplyr)
df <- df %>%
  select(age, sex, moa, mod, los, where(~!any(is.na(.))))
```

```{r}
library(tidyr)
drop_na(df)
```

6. Which variables records the month of admission, month of discharge?
```{r}
table(df$moa) #month of admission
table(df$mod) #month of discharge 
```

7. Which month admitted the most number of patients? Which month admitted the most number of male patients?
```{r}
#The month that admitted the most number of patients is March.
df %>%
  group_by(moa) %>%
  summarise(total_patients = n()) %>%
  arrange(-total_patients)
```

```{r}
#The month that admitted the most number of male patients is March.
df %>%
  group_by(moa) %>%
  summarise(malePatients = sum(sex == "1")) %>%
  arrange(-malePatients)
```

8. Which month has the most number of teenage female patients?
```{r}
#The month that has the most number of teenage female patients is March.

df$teenAges <- case_when(df$age <13 ~ 'child', df$age<=19 ~'teenager', TRUE ~ 'adult')

df %>%
  filter(sex == 2, teenAges == 'teenager') %>%
  group_by(moa) %>%
  summarise(teenageFemales = n()) %>%
  arrange(-teenageFemales)
```
9. Which provider has the most number of female patients in October?
```{r}
#Rhode Island Hospital (7205) has the most number of female patients in October.
df %>%
  filter(sex == 2, moa == 10) %>%
  group_by(provider) %>%
  summarise(femalePatients = n()) %>%
  arrange(-femalePatients)
```

10. Are female patients older than male patients on average?

```{r}
#Female patients are older than male patients on average. 

df %>%
  group_by(sex) %>%
  summarise(averageAge = mean(age, na.rm = TRUE)) %>%
  arrange(-averageAge)
```


11. Calculate the average age of patients by months. Which month has the oldest patients on average?

```{r}
#December has the oldest patients on average. 

df %>%
  group_by(moa)%>%
  summarise(averageAge = mean(age, na.rm= TRUE)) %>%
  arrange(-averageAge)
```


12. What is the name of the provider who has the highest total charge?

```{r}
#The provider who has the highest total charge is Rhode Island Hospital (7205).

df %>%
  group_by(provider) %>%
  summarise(highestTotalCharge = sum(tot)) %>%
  arrange(-highestTotalCharge)
```


13. What is the name of the provider that has the least total charge for teenage male on average?

```{r}
#Newport Hospital (7201) has the least total charge for teenage male patients on average.

df %>%
  filter(teenAges == "teenager", sex == 1) %>%
  group_by(provider) %>%
  summarise(avgTotalCharge = mean(tot, na.rm = TRUE)) %>%
  arrange(avgTotalCharge)
  
```


14. Create a season (Spring, Summer, Fall, Winter) variable. Calculate the length of stays by season. Which season has the longest length of stays on average?

```{r}
df <- df %>%
  mutate(season = case_when(
    moa %in% c(12, 1, 2) ~ "Winter",
    moa %in% c(3, 4, 5) ~ "Spring",
    moa %in% c(6, 7, 8) ~ "Summer",
    moa %in% c(9, 10, 11) ~ "Fall"))

#Winter has the longest length of stays on average.
df %>%
  group_by(season) %>%
  summarise(avgLOS = mean(los, na.rm = TRUE)) %>%
  arrange(-avgLOS)
```


15. On average, how much a 20 year-old male get charged for staying 1 day in the Fall season?

```{r}
#The average charge for a 20 yeaer old male in the Fall is $10,697.54.
df %>%
  filter(sex == 1, age == 20, season == "Fall", los == 1) %>%
  summarise(averageCharge = mean(tot, na.rm = TRUE))
```


16. Write a paragraph to summarize the section and give your comments on the results. You could do some other calculations to support your points.

